A comparison of model-based and design-based approaches for spatial data.

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated introduction.

Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.

<!-- ????????? -->
We define spatial data as [blah blah blah, any data collected at specific locations blah blah blah blah]. Examples include [do we want to give a list of general examples?]. 
<!-- ???????????? -->

@de1990model give an early comparison of these approaches, quashing the belief that design-based approaches could not be used for spatially correlated data. Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial]. More recent overviews include @brus2020statistical and @wang2012spatial, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.

>>>>

Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data. We define spatial data as variables measured at specific geographic locations. @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data. Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial]. More recent overviews include @brus2020statistical and @wang2012spatial, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated introduction.

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data.

We define spatial data as {[}blah blah blah, any data collected at
specific locations blah blah blah blah{]}. Examples include {[}do we
want to give a list of general examples?{]}.

De Gruijter and Ter Braak (1990) give an early comparison of these
approaches, quashing the belief that design-based approaches could not
be used for spatially correlated data. Thereafter, several comparisons
between design-based and model-based for spatial data have been
considered, but they tend to compare design-based approaches that ignore
spatial location in sampling to model-based approaches (Brus and De
Gruijter, 1997; Brus, 2020; Ver Hoef, 2002; Ver Hoef, 2008). No
comparison has been made between design-based approaches that
incorporate spatial location in sampling and model-based approaches.

wang2012spatial - sampling and inference steps, thorough overview in
terms of spatial context, lots of references

In Section \ref{sec:background}, we compare sampling and estimation
procedures between the design-based approach and the model-based
approach. In Section \ref{sec:numstudy}, we use simulated and real data
to compare the properties of parameter estimates from both aproaches.
And in Section \ref{sec:discussion}, we end with a discussion and
provide directions for future research.

@@

>>>>

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as variables measured at specific
geographic locations. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
location in sampling to model-based approaches (Brus and De Gruijter,
1997; Ver Hoef, 2002; Ver Hoef, 2008). More recent overviews include
Brus (2020) and (\textbf{wang2012spatial?}), but no numerical comparison
has been made between design-based approaches that incorporate spatial
location in sampling and model-based approaches.

The rest of this paper is organized as follows. In Section
\ref{sec:background}, we compare sampling and estimation procedures
between the design-based approach and the model-based approach. In
Section \ref{sec:numstudy}, we use simulated and real data to compare
the properties of parameter estimates from both aproaches. And in
Section \ref{sec:discussion}, we end with a discussion and provide
directions for future research.

@@

\hypertarget{applied-example}{%
\subsection{Applied Example}\label{applied-example}}
